---
title: "Reproducible Research: Peer Assessment 1"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

Load libraries and set code chunk defaults.

```{r setup}
library(readr)
library(ggplot2)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

Load the data and display the first 6 rows.

```{r results = 'markup'}
file <- "activity.zip"
df <- read_csv(file)
print(head(df))
```

## What is mean total number of steps taken per day?

```{r histogram-daily-steps}
daily <- df %>%
  na.omit() %>%
  group_by(date) %>%
  summarize(total = sum(steps))

ggplot(data=daily, aes(daily$total)) +
  geom_histogram(binwidth = 1000, fill="steelblue") +
  theme_minimal() +
  labs(title = "Histogram of Total Daily Steps", x = "Total Daily Steps", y = "Count")
```

```{r}
ave.steps <- format(mean(daily$total), scientific = FALSE)
median.steps <- format(median(daily$total), scientific = FALSE)
```

The mean total number of steps taken each day is `r ave.steps` and the median total number of steps taken per day is `r median.steps`.


## What is the average daily activity pattern?
```{r}
interval.ave <- df %>%
  na.omit() %>%
  group_by(interval) %>%
  summarize(ave = mean(steps))

ggplot(data = interval.ave, aes(x = interval, y = ave))+
  geom_line(color = "#00AFBB", size = 2) +
  theme_minimal() +
  labs(title = "Time Series of Ave Steps per Interval", x = "Interval", y = "Ave Steps per Interval")
```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
