---
title: "Reproducible Research: Peer Assessment 1"
output:
  html_document:
    keep_md: true
---

Load libraries and set code chunk defaults.

```{r setup}
library(readr)
library(dplyr)
library(ggplot2)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

Load the data and display the first 6 rows.

```{r results = 'asis'}
file <- "activity.zip"
df <- read_csv(file)
print(head(df))
```

## What is mean total number of steps taken per day?

```{r histogram-daily-steps}
daily <- df %>%
  na.omit() %>%
  group_by(date) %>%
  summarize(total = sum(steps))

ggplot(data=daily, aes(daily$total)) +
  geom_histogram(binwidth = 1000, fill="steelblue") +
  theme_minimal() +
  labs(title = "Histogram of Total Daily Steps", x = "Total Daily Steps")
```

```{r}
ave.steps <- format(mean(daily$total), scientific = FALSE)
median.steps <- format(median(daily$total), scientific = FALSE)
```

The mean total number of steps taken each day is `r ave.steps` and the median total number of steps taken per day is `r median.steps`.


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
